% !TeX program = lualatex
% =====================================================================
% 04-functions.tex --- Three complete function studies, eight steps each.
%
% Each study follows the eight-step method (Müller). The table at step 6
% carries three lines -- f'' (convexity), f' (growth), f (variation) --
% all from one shared <fn> object; the plot at step 8 comes from the same
% object. Sections in ROMAN, steps as plain numbers.
%
% The f'' line is optional in scholatex (attribute second:). It is shown
% here because convexity is one of the eight steps; for lower grades it
% would simply be omitted.
% =====================================================================
\documentclass[margins=15, size=12, lang=en]{scholatex}
\begin{document}

let title = <Navy b 18pt c>
let study = <Navy b section ROMAN>
let step  = <Navy b subsection roman >
let p     = <tab>

<title>Function studies

% ====================================================================
<line>
<study title:{Polynomial: $f(x) = -x^4 + 2x^2 + 1$}>{
<p>An even quartic.
}

let f = <fn name:{f(x)}
            expr:{-x^4 + 2x^2 + 1}
            x:{-inf | -1 | -1/sqrt(3) | 0 | 1/sqrt(3) | 1 | +inf}
            second:{- | - | + | + | - | -}
            deriv:{+ | + | + | - | - | -}
            var:{-inf / 2 \ 1 / 2 \ -inf}>

<step title:{Domain}>{
<p>$f$ is a polynomial: its domain is all of $R$.
}

<step title:{Parity}>{
<p>$f(-x) = -x^4 + 2x^2 + 1 = f(x)$, so $f$ is even; the graph is
symmetric about the $y$-axis.
}

<step title:{Sign}>{
<p>$f(0) = 1 > 0$ and $f$ tends to $-inf$ at both ends, so $f$ vanishes
at two symmetric points and is positive between them, negative outside.
}

<step title:{Vertical asymptotes}>{
<p>None: $f$ is defined and continuous on all of $R$.
}

<step title:{Affine asymptotes}>{
<p>None: a quartic grows faster than any line, the ratio of $f(x)$ to $x$ tends to $+- inf$. There is no affine asymptote.
}

<step title:{Growth and critical points}>{
<p>$f'(x) = -4x^3 + 4x = 4x(1 - x^2)$ vanishes at $-1$, $0$, $1$: two
maxima $f(-1) = f(1) = 2$ and a local minimum $f(0) = 1$.
}

<step title:{Convexity and inflection points}>{
<p>$f''(x) = -12x^2 + 4 = 4(1 - 3x^2)$ vanishes at $+- 1/sqrt(3)$:
$f$ is convex between them and concave outside, with two inflection
points. The table gathers $f''$, $f'$ and $f$.
}

<vartab f>

<step title:{Graph}>{
<plot f samples:200 x:{-2, 2} y:{-3, 3}>
}

% ====================================================================
<study title:{Rational, horizontal asymptote: $g(x) = x^2/(x^2-2x+2)$}>{
<p>Denominator without real root.
}

let g = <fn name:{g(x)}
            expr:{x^2/(x^2 - 2x + 2)}
            x:{-inf | 1-sqrt(3) | 0 | 1 | 2 | 1+sqrt(3) | +inf}
            second:{- | + | + | - | - | +}
            deriv:{- | - | + | + | - | -}
            var:{1 \ 0 / 2 \ 1}>

<step title:{Domain}>{
<p>The discriminant of $x^2 - 2x + 2$ is $-4 < 0$, so the denominator
never vanishes: $g$ is defined on $R$.
}

<step title:{Parity}>{
<p>$g(-x) != g(x)$ and $g(-x) != -g(x)$: $g$ is neither even nor
odd.
}

<step title:{Sign}>{
<p>$g(x) = x^2 / (x^2 - 2x + 2)$ is a ratio of a square by a positive
quantity, so $g(x) >= 0$, vanishing only at $x = 0$.
}

<step title:{Vertical asymptotes}>{
<p>None: the denominator never vanishes.
}

<step title:{Horizontal asymptote}>{
<p>$g(x)$ tends to $1$ as $x$ tends to $+- inf$, so the line $y = 1$
is a horizontal asymptote on both sides.
}

<step title:{Growth and critical points}>{
<p>$g'(x) = 2x(2-x)/(x^2-2x+2)^2$ vanishes at $0$ and $2$: a minimum
$g(0) = 0$ and a maximum $g(2) = 2$.
}

<step title:{Convexity and inflection points}>{
<p>$g''$ vanishes at $1-sqrt(3)$, $1$ and $1+sqrt(3)$, giving three
inflection points. The table gathers $g''$, $g'$ and $g$.
}

<vartab g>

<step title:{Graph}>{
<plot g samples:200 x:{-6, 6} y:{-1, 3}>
}

% ====================================================================
<study title:{Rational, vertical asymptote: $k(x) = (x^2+1)/(x-1)$}>{
<p>A pole at $x = 1$ and an affine asymptote.
}

let k = <fn name:{k(x)}
            expr:{(x^2+1)/(x-1)}
            x:{-inf | 1-sqrt(2) | 1 | 1+sqrt(2) | +inf}
            second:{- | - || + | +}
            deriv:{+ | - || - | +}
            var:{-inf / 2-2sqrt(2) \ -inf || +inf \ 2+2sqrt(2) / +inf}>

<step title:{Domain}>{
<p>The denominator $x - 1$ vanishes at $x = 1$: the domain is
$]-inf, 1[$ union $]1, +inf[$.
}

<step title:{Parity}>{
<p>The domain is not centred at $0$, so $k$ is neither even nor odd.
}

<step title:{Sign}>{
<p>$x^2 + 1 > 0$ always, so $k(x)$ has the sign of $x - 1$: negative on
$]-inf, 1[$, positive on $]1, +inf[$.
}

<step title:{Vertical asymptote}>{
<p>At $x = 1$, $k(x)$ tends to $-inf$ on the left and $+inf$ on the
right: the line $x = 1$ is a vertical asymptote.
}

<step title:{Affine asymptote}>{
<p>Dividing gives $k(x) = x + 1 + 2/(x-1)$, so $k(x) - (x+1)$ tends to
$0$: the line $y = x + 1$ is an affine asymptote.
}

<step title:{Growth and critical points}>{
<p>$k'(x) = (x^2-2x-1)/(x-1)^2$ vanishes at $1 +- sqrt(2)$: a maximum
$k(1-sqrt(2)) = 2 - 2sqrt(2)$ and a minimum $k(1+sqrt(2)) = 2 + 2sqrt(2)$.
}

<step title:{Convexity}>{
<p>$k''(x) = 4/(x-1)^3$ never vanishes but changes sign at the pole:
$k$ is concave on $]-inf, 1[$ and convex on $]1, +inf[$, with no
inflection point. The table gathers $k''$, $k'$ and $k$.
}

<vartab k>

<step title:{Graph}>{
<plot k samples:200 x:{-4, 6} y:{-10, 12}>
}

\end{document}
